They're Clutching up! Team Momentum in Round-Based Esports

### 2022-08-14 paper/paper.Rmd
data paths
add init streak blocks of code

cod_rounds <- qs::qread('cod_rounds.qs')
```

@@

>>>>

cod_rounds <- qs::qread('data/cod_rounds.qs')
```

@@

# Momentum

```{r}
#| label: qs-funcs
#| eval: true
load_qs <- function(name) {
  res <- qs::qread(sprintf('../data/%s.qs', name))
  assign(name, value = res, envir = .GlobalEnv)
}
save_qs <- function(df, name = deparse(substitute(df))) {
  qs::qsave(df, sprintf('../data/%s.qs', name))
}
```

```{r}
#| label: cod_round_win_prop_after_x_wins
## http://keyonvafa.com/hot-hand/
get_post_streak_prob <- function(n, k, p = 0.5) {
  tosses <- rbinom(n, 1, p)
  runs <- rle(tosses)
  n_neg_after <- length(which(runs$values == 1 & runs$lengths >= k))
  n_pos_after <- sum(runs$lengths[which(runs$values == 1 & runs$lengths >= k)] - k)

  ## edge case
  if (n %in% cumsum(runs$lengths)[which(runs$values == 1 & runs$lengths >= k)]) {
    n_neg_after <- n_neg_after - 1
  }

  n_pos_after / (n_pos_after + n_neg_after)
}

simulate_post_streak_prob <- function(sims = 10000, seed = 42, ...) {
  withr::local_seed(seed)
  rerun(
    sims,
    get_post_streak_prob(...)
  ) |> 
    flatten_dbl() |> 
    mean(na.rm = TRUE)
}

cod_round_streaks <- cod_rounds |> 
  group_by(series_id, team) |> 
  mutate(
    won_prior_round1 = lag(win_round, n = 1, default = NA),
    won_prior_round2 = lag(win_round, n = 2, default = NA),
    won_prior_round3 = lag(win_round, n = 3, default = NA),
    won_prior_round4 = lag(win_round, n = 4, default = NA),
    won_prior_round5 = lag(win_round, n = 5, default = NA)
  ) |> 
  ungroup() |>
  select(
    series_id,
    team,
    is_offense,
    round,
    n_rounds,
    win_round,
    starts_with('won_prior_round')
  )

.postprocess_round_streaks <- function(k, round_streaks, probs_nx_k, ...) {

  cols <- sprintf('won_prior_round%d', 2:k)
  col_syms <- syms(cols)
  round_streaks_after_x <- round_streaks |> 
    drop_na(!!!col_syms) |> 
    count(n_rounds, ..., win_round, won_prior_round1, !!!col_syms, sort = TRUE)

  round_win_prop_after_x_wins <- round_streaks_after_x |> 
    filter(won_prior_round1, !!!col_syms)

  suppressMessages(
    round_win_prop_after_x_wins_w_probs <- round_win_prop_after_x_wins |> 
      group_by(n_rounds, ...) |> 
      mutate(
        total = sum(n),
        prop = n / total
      ) |> 
      ungroup() |> 
      filter(win_round) |> 
      select(n_rounds, ..., n, total, prop) |> 
      arrange(n_rounds, ...) |> 
      inner_join(
        probs_nx_k
      )
  )

  round_win_prop_after_x_wins_w_probs |> 
    nest(data = -c(n_rounds, ...)) |> 
    mutate(
      gt_p_value = map_dbl(
        data, 
        ~binom.test(.x$n, .x$total, .x$expected_poststreak_p, alternative = 'greater') |> 
          broom::tidy() |>
          pull(p.value)
      ),
      lt_p_value = map_dbl(
        data, 
        ~binom.test(.x$n, .x$total, .x$expected_poststreak_p, alternative = 'less') |> 
          broom::tidy() |>
          pull(p.value)
      ),
      gt_is_signif = gt_p_value <= 0.05,
      lt_is_signif = lt_p_value <= 0.05,
      is_signif = case_when(
        gt_is_signif ~ 'Better',
        lt_is_signif ~ 'Worse',
        TRUE ~ 'Neither'
      )
    ) |> 
    unnest(data)
}

summarize_cod_win_prop_after_x_wins <- function(k) {
  probs_nx_k <- tibble(
    n_rounds = 7L:11L
  ) |> 
    mutate(
      prior_p = 6L / n_rounds,
      expected_poststreak_p = map2_dbl(
        n_rounds, prior_p, 
        ~simulate_post_streak_prob(n = ..1, k = !!k, p = ..2)
      )
    )

  .postprocess_round_streaks(
    k = k,
    round_streaks = cod_round_streaks,
    probs_nx_k = probs_nx_k
  )
}

cod_round_win_prop_after_x_wins <- tibble(win_streak = 2L:4L) |>
  mutate(
    data = map(win_streak, summarize_cod_win_prop_after_x_wins)
  ) |>
  unnest(data)
save_qs(cod_round_win_prop_after_x_wins)
```

```{r}
#| label: cod_round_win_prop_after_x_wins_and_rounds
summarize_cod_win_prop_after_x_wins_y_rounds <- function(k) {
  probs_nx_k <- crossing(
    n_rounds = 7L:11L,
    round = 3L:11L
  ) |> 
    filter(round >= (k + 1L)) |> 
    mutate(
      prior_p = 6L / n_rounds,
      expected_poststreak_p = map2_dbl(
        n_rounds, prior_p, 
        ~simulate_post_streak_prob(n = ..1, k = !!k, p = ..2)
      )
    )

  .postprocess_round_streaks(
    k = k,
    round_streaks = cod_round_streaks,
    probs_nx_k = probs_nx_k,
    round
  )
}

cod_round_win_prop_after_x_wins_y_rounds <- tibble(win_streak = 2L:4L) |>
  mutate(
    data = map(win_streak, summarize_cod_win_prop_after_x_wins_y_rounds)
  ) |>
  unnest(data)
save_qs(cod_round_win_prop_after_x_wins_y_rounds)
```

Despite the implication that one can model team round win probability in CoD SnD as a constant, anecdotally many believe in momentum and positive recency, i.e. the "hot hand".

Miller et al. [-@miller2018]
